\Large \bf Uncovering Bugs in Distributed Storage Systems\\during Testing (not in Production!)

### 2015-12-25 vnext.tex
major update to address PC and shepherd comments

To uncover the elusive extent repair bug in \Azure Storage vNext, its developers wrote a test harness using \psharp. The developers expected that it was more likely for the bug to occur in the ExtMgr logic, rather than in the EN logic. Hence, they focused on testing the real ExtMgr using modeled ENs. The test harness for vNext consists of the following \psharp machines (as shown in Figure~\ref{fig:azurestoremodel}):
\begin{description}
\item[ExtentManager] acts as a thin wrapper machine for the real ExtMgr component in vNext (see \S\ref{sec:method:wrap_target}).

\item[ExtentNode] is a simple model of an EN (see \S\ref{sec:method:mock_en}).

@@

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{img/mocked_vnext}
\caption{Real Extent Manager with a modeled environment (each box represents one \psharp machine).}
\label{fig:azurestoremodel}
\vspace{-2mm}
\end{figure}

@@

  void Init() {
    ExtMgr = new ExtentManager();
    ExtMgr.NetEngine = new MockedNetEngine(); // mock network
    ExtMgr.IsMockingTimer = true;	 // disable internal timer
  }

@@

  [OnEvent(TimerTickEvent, ProcessExtentRepair)]
  void ProcessExtentRepair() {
    // extent repair loop driven by external timer
    ExtMgr.ProcessEvent(new ExtentRepairEvent());
  }
}
\end{lstlisting}
\vspace{-3mm}
\caption{The real Extent Manager is wrapped inside the \texttt{ExtentManager} \psharp machine.}
\label{fig:wrap_target}
%\vspace{-2mm}
\end{figure}

@@

\begin{figure}[t]
\centering
\includegraphics[width=.95\linewidth]{img/extent_manager}
\caption{Internal components of the real Extent Manager in \Microsoft \Azure Storage vNext.}
\label{fig:extentmanager}
\vspace{-5mm}
\end{figure}

\textbf{Internals of the real Extent Manager}.
The real ExtMgr (see Figure~\ref{fig:extentmanager}) contains two data structures related to extent replication and repair: \texttt{ExtentCenter} and \texttt{ExtentNodeMap}. The \texttt{ExtentCenter} data structure maintains the records that map from extents to their hosting ENs. It is updated upon receiving the periodic sync reports from the ENs. Recall that the sync report from a particular EN lists all the extents stored at the EN. Its purpose is to update ExtMgr's possible out-of-date view of the EN with the ground truth. \texttt{ExtentNodeMap} records the latest heartbeat time from every EN.

@@

\textbf{Intercepting network messages}.
The real ExtMgr uses a network engine to asynchronously send messages to the ENs. The \psharp test harness mocks the original network engine in vNext by overriding its interface. This enables the mocked network engine (we use the terms \textit{mocked} and \textit{modeled} interchangeably) to intercept all outbound messages and relay them to the \texttt{TestingDriver} machine, which is responsible for dispatching the messages to the corresponding \texttt{ExtentNode} machines. As shown in Figure~\ref{fig:enginecode}, the mocked network engine intercepts the outbound messages from the ExtMgr and invokes \texttt{PSharp.Send(...)} to asynchronously relay the messages to \texttt{TestingDriver}. This mocked network engine replaces the real network engine in the wrapped ExtMgr, as shown in Figure~\ref{fig:wrap_target}.

@@

// mocked engine for intercepting Extent Manager messages
class MockedNetEngine : NetworkEngine {
  public override void SendMessage(Socket s, Message msg) {
    // intercept and relay Extent Manager messages
    PSharp.Send(TestingDriver, new ExtMgrMsgEvent(), s, msg);
  }
}
\end{lstlisting}
\vspace{-4mm}
\caption{Mocked network engine in vNext.}
\label{fig:enginecode}
\vspace{-2mm}
\end{figure}

Intercepting all network messages and dispatching them through the \psharp runtime is important for two reasons. First, it allows \psharp to systematically explore the interleavings between asynchronous event handlers in the system. Second, the mocked network engine could leverage the support for controlled nondeterministic choices in \psharp, and choose to drop the messages in a non-deterministic fashion, in case emulating message loss is desirable (not shown in this example).

Messages coming from the \texttt{ExtentNode} machines do \emph{not} go through the mocked network engine. They are delivered to the \texttt{ExtentManager} machine directly and trigger an action that invokes the messages on the wrapped ExtMgr with \texttt{ExtMgr.ProcessMessage} (see Figure~\ref{fig:wrap_target}). The benefit of this approach is that the real ExtMgr can be tested without modifying its code; the ExtMgr is simply unaware of the \psharp test harness and behaves as if it is running in a real distributed environment and communicating with real ENs.

\subsection{The ExtentNode machine}
\label{sec:method:mock_en}

@@

  // extent repair logic
  ...
  [OnEvent(ExtentCopyResponseEvent, ProcessCopyResponse)]
  void ProcessCopyResponse(ExtentCopyResponse response) {
    // extent copy response from source replica
    if (IsCopySucceeded(response)) {
      var rec = GetExtentRecord(response);
      ExtCtr.AddOrUpdate(rec); // update ExtentCenter
    }
  }

@@

  // extent node failure logic
  [OnEvent(FailureEvent, ProcessFailure)]
  void ProcessFailure() {
    // notifies the monitor that this EN failed
    PSharp.Notify<RepairMonitor>(new ENFailedEvent(), this);
    PSharp.Halt(); // terminate this P# machine
  }
}
\end{lstlisting}
\vspace{-4mm}
\caption{The modeled EN in vNext.}
\label{fig:mocked_en}
\vspace{-2mm}
\end{figure}

@@

The \psharp test harness leverages components of the real vNext system whenever it is appropriate. For example, \texttt{ExtentNode} re-uses the \texttt{ExtentCenter} data structure, which is used inside a real EN for extent bookkeeping. In the modeled extent repair logic, \texttt{ExtentNode} takes action upon receiving an extent repair request from the \texttt{ExtentManager} machine. It sends a copy request to a source \texttt{ExtentNode} machine where a replica is stored. After receiving an \texttt{ExtentCopyRespEvent} event from the source machine, it updates the \texttt{ExtentCenter}, as illustrated in Figure~\ref{fig:mocked_en}.

In the modeled EN sync logic, the machine is driven by an external timer modeled in \psharp (see \S\ref{sec:method:timer}). It prepares a sync report with \texttt{extCtr.GetSyncReport(...)}, and then asynchronously sends the report to \texttt{ExtentManager} using \texttt{PSharp.Send(...)}. The \texttt{ExtentNode} machine also includes failure-related logic, which is discussed in detail in \S\ref{sec:method:driver}.

@@

  [OnEvent(RepeatedEvent, GenerateTimerTick)]
  void GenerateTimerTick() {
    // nondeterministic choice controlled by P#
    if (PSharp.Nondet())
      PSharp.Send(Target, new TimerTickEvent());
    PSharp.Send(this, new RepeatedEvent()); // loop
  }
}
\end{lstlisting}
\vspace{-4mm}
\caption{Modeling timer expiration using \psharp.}
\label{fig:mocked_timer}
\vspace{-2mm}
\end{figure}

@@

Figure~\ref{fig:mocked_timer} shows the \texttt{Timer} machine in the test harness. \texttt{Timer} invokes the \psharp method \texttt{Nondet()}, which generates a nondeterministic choice controlled by the \psharp runtime. Using \texttt{Nondet()} allows the machine to nondeterministically send a timeout event to its target (the \texttt{ExtentManager} or \texttt{ExtentNode} machines). The \psharp testing engine has the freedom to schedule arbitrary interleavings between these timeout events and all other regular system events.

@@

  // other test harness logic
  ...
  void InjectNodeFailure() {
    // nondeterministically choose an EN using P#
    var node = (Machine)PSharp.Nondet(ExtentNodes);    
    PSharp.Send(node, new FailureEvent()); // fail chosen EN
  }
}
\end{lstlisting}
\vspace{-4mm}
\caption{The TestingDriver machine in vNext.}
\label{fig:testing_driver}
\vspace{-2mm}
\end{figure}

@@

Figure~\ref{fig:testing_driver} illustrates how \texttt{TestingDriver} leverages \psharp to inject nondeterministic failures. It uses \texttt{Nondet()} to nondeterministically choose an \texttt{ExtentNode} machine, and then sends a \texttt{FailureEvent} to the chosen machine to emulate an EN failure. As shown in the earlier Figure~\ref{fig:mocked_en}, the chosen \texttt{ExtentNode} machine processes the \texttt{FailureEvent}, notifies the liveness monitor of its failure (see \S\ref{sec:method:monitor}) and terminates itself by invoking the \psharp method \texttt{Halt()}. \psharp not only enumerates interleavings of asynchronous event handlers, but also the values returned by calls to \texttt{Nondet()}, thus enumerating different failure scenarios.

@@

\texttt{RepairMonitor} is a \psharp liveness monitor (see \S\ref{sec:bg:bugs}) that transitions between a cold and a hot state. Whenever an EN fails, the monitor is notified with an \texttt{ENFailedEvent} event. As soon as the number of extent replicas falls below a specified target (three replicas in the current \psharp test harness), the monitor transitions into the hot \emph{repairing} state, waiting for all missing replicas to be repaired. Whenever an extent replica is repaired, \texttt{RepairMonitor} is notified with an \texttt{ExtentRepairedEvent} event. When the replica number reaches again the target, the monitor transitions into the cold \emph{repaired} state, as illustrated in the code snippet of Figure~\ref{fig:monitor}.

@@

\subsection{Liveness Bug in \Azure Storage vNext}
\label{sec:method:azurestore}

>>>>

Microsoft Azure Storage is a cloud storage system that provides customers the ability to store seemingly limitless amounts of data. It has grown from tens of petabytes in 2010 to exabytes in 2015, with the total number of objects stored well-exceeding 60 trillion.
%~\cite{greenberg2015keynote}.

Azure Storage vNext is the next generation storage system for Microsoft Azure, where the primary design target is to scale the storage capacity by more than 100$\times$. Similar to the current system, vNext employs containers, called \emph{extents}, to store data. Extents can be several gigabytes each, consisting of many data blocks, and are replicated over multiple \emph{Extent Nodes} (ENs). However, in contrast to the current system, which uses a Paxos-based, centralized mapping from extents to ENs, vNext achieves scalability by using a \emph{distributed mapping}. In vNext, extents are divided into partitions, with each partition managed by a light-weight \emph{Extent Manager} (ExtMgr). This partitioning is illustrated in Figure~\ref{fig:vnext}.

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{img/azurestore}
\vspace{-5mm}
\caption{Top-level components for extent management in Microsoft Azure Storage vNext.}
\label{fig:vnext}
\vspace{-2mm}
\end{figure}

One of the many responsibilities of an ExtMgr is to ensure that every extent maintains enough \emph{replicas} in the system. To achieve this, an ExtMgr receives frequent periodic \emph{heartbeat} messages from every EN. The failure of an EN is detected by missing heartbeats. An ExtMgr also receives less frequent, but still periodic, \emph{synchronization reports} from every EN. The sync reports list all the extents (and associated metadata) stored on the EN. Based on these two types of messages, an ExtMgr identifies which ENs have failed, and which extents are affected by the failure and are missing replicas as a result. The ExtMgr then schedules tasks to repair the affected extents and distributes the tasks to the necessary ENs. The ENs then repair the extents from their existing replicas in the system and lazily update the ExtMgr via their next periodic sync reports. All this communication between an ExtMgr and the ENs occurs via network engines installed in each component of vNext (see Figure~\ref{fig:vnext}).

To ensure correctness, the developers of vNext have instrumented extensive, multiple levels of testing:
\begin{enumerate}
\item \emph{Unit testing}, in which emulated heartbeats and sync reports are sent to an ExtMgr. These tests check that the messages are processed as expected.

\item \emph{Integration testing}, in which an ExtMgr is launched together with multiple ENs. An EN failure is subsequently injected. These tests check that the affected extents are eventually repaired.

\item \emph{Stress testing}, in which an ExtMgr is launched together with multiple ENs and multiple extents. The test keeps repeating the following process: injecting an EN failure, launching a new EN and checking that the affected extents are eventually repaired.
\end{enumerate}

\noindent
Despite of the extensive testing efforts, the vNext developers have been plagued for months by an elusive bug in the ExtMgr logic. All the unit test suites and integration test suites successfully pass every single time. However, the stress test suite fails \emph{from time to time} after very long executions; in these cases, certain replicas of some extents fail and are never repaired.

This bug has proven to be difficult to identify, reproduce and troubleshoot. First, an extent never being repaired is \emph{not} a property that can be easily checked. Second, the bug seems to manifest only in very long executions. Finally, by the time that the bug is detected, very long execution traces have been collected, which makes manual inspection tedious and ineffective.
%To uncover this bug, as well as many other tricky bugs, the vNext developers are in constant search of a generic and systematic approach for testing distributed systems.

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{img/modeled_vnext}
\vspace{-5mm}
\caption{Real Extent Manager with a modeled environment (each box represents one \psharp machine).}
\label{fig:azurestoremodel}
\vspace{-2mm}
\end{figure}

To uncover the elusive extent repair bug in Azure Storage vNext, its developers wrote a test harness using \psharp. The developers expected that it was more likely for the bug to occur in the ExtMgr logic, rather than in the EN logic. Hence, they focused on testing the real ExtMgr using modeled ENs. The test harness for vNext consists of the following \psharp machines (as shown in Figure~\ref{fig:azurestoremodel}):

\begin{description}
\item[ExtentManager] acts as a thin wrapper machine for the real ExtMgr component in vNext (see \S\ref{sec:method:wrap_target}).

\item[ExtentNode] is a simple model of an EN (see \S\ref{sec:method:model_en}).

@@

@@

  void Init() {
    ExtMgr = new ExtentManager();
    ExtMgr.NetEngine = new ModelNetEngine(); // model network
    ExtMgr.IsModelingTimer = true;	 // disable internal timer
  }

@@

  [OnEvent(TimerTickEvent, ProcessExtentRepair)]
  void ProcessExtentRepair() {
    // extent repair loop driven by external timer
    ExtMgr.ProcessEvent(new ExtentRepairEvent());
  }
}
\end{lstlisting}
\vspace{-2mm}
\caption{The real Extent Manager is wrapped inside the \texttt{ExtentManager} \psharp machine.}
\label{fig:wrap_target}
\end{figure}

@@

\begin{figure}[t]
\centering
\includegraphics[width=.95\linewidth]{img/extent_manager}
\vspace{-2mm}
\caption{Internal components of the real Extent Manager in Microsoft Azure Storage vNext.}
\label{fig:extentmanager}
\end{figure}

\textbf{Real Extent Manager}.
The real ExtMgr (see Figure~\ref{fig:extentmanager}) contains two data structures: \texttt{ExtentCenter} and \texttt{ExtentNodeMap}. The \texttt{ExtentCenter} maintains the records that map from extents to their hosting ENs. It is updated upon receiving the periodic sync reports from the ENs. Recall that the sync report from a particular EN lists all the extents stored at the EN. Its purpose is to update ExtMgr's possible out-of-date view of the EN with the ground truth. The \texttt{ExtentNodeMap} records the latest heartbeat time from every EN.

@@

\textbf{Intercepting network messages}.
The real ExtMgr uses a network engine to asynchronously send messages to the \texttt{ExtentNode} machines. The \psharp test harness models the original network engine in vNext by overriding its interface. The modeled network engine (see Figure~\ref{fig:enginecode}) intercepts all outbound messages from the ExtMgr, and invokes \texttt{PSharp.Send(...)} to asynchronously relay the messages to \texttt{TestingDriver} machine, which is responsible for dispatching the messages to the corresponding ENs. This modeled network engine replaces the real network engine in the wrapped ExtMgr (see Figure~\ref{fig:wrap_target}.

@@

// modeled engine for intercepting Extent Manager messages
class ModelNetEngine : NetworkEngine {
  public override void SendMessage(Socket s, Message msg) {
    // intercept and relay Extent Manager messages
    PSharp.Send(TestingDriver, new ExtMgrMsgEvent(), s, msg);
  }
}
\end{lstlisting}
\vspace{-4mm}
\caption{Modeled network engine in vNext.}
\label{fig:enginecode}
\vspace{-3mm}
\end{figure}

Intercepting all network messages and dispatching them through \psharp is important for two reasons. First, it allows \psharp to systematically explore the interleavings between asynchronous event handlers in the system. Second, the modeled network engine could leverage the support for controlled nondeterministic choices in \psharp, and choose to drop the messages in a non-deterministic fashion, in case emulating message loss is desirable (not shown in this example).

Messages coming from the \texttt{ExtentNode} machines do \emph{not} go through the modeled network engine. They are delivered to the \texttt{ExtentManager} machine directly and trigger an action that invokes the messages on the wrapped ExtMgr with \texttt{ExtMgr.ProcessMessage} (see Figure~\ref{fig:wrap_target}). The benefit of this approach is that the real ExtMgr can be tested without modifying its code; the ExtMgr is simply unaware of the \psharp test harness and behaves as if it is running in a real distributed environment and communicating with real ENs.

\subsection{The ExtentNode machine}
\label{sec:method:model_en}

@@

  [OnEvent(ExtentCopyResponseEvent, ProcessCopyResponse)]
  void ProcessCopyResponse(ExtentCopyResponse response) {
    // extent copy response from source replica
    if (IsCopySucceeded(response)) {
      var rec = GetExtentRecord(response);
      ExtCtr.AddOrUpdate(rec); // update ExtentCenter
    }
  }

@@

  // extent node failure logic
  [OnEvent(FailureEvent, ProcessFailure)]
  void ProcessFailure() {
    // notifies the monitor that this EN failed
    PSharp.Notify<RepairMonitor>(new ENFailedEvent(), this);
    PSharp.Halt(); // terminate this P# machine
  }
}
\end{lstlisting}
\vspace{-4mm}
\caption{Code snippet of the modeled EN in vNext.}
\label{fig:modeled_en}
\vspace{-2mm}
\end{figure}

@@

The \psharp test harness leverages components of the real vNext system whenever it is appropriate. For example, \texttt{ExtentNode} re-uses the \texttt{ExtentCenter} data structure, which is used inside a real EN for extent bookkeeping. In the modeled extent repair logic, \texttt{ExtentNode} takes action upon receiving an extent repair request from the \texttt{ExtentManager} machine. It sends a copy request to a source \texttt{ExtentNode} machine where a replica is stored. After receiving an \texttt{ExtentCopyRespEvent} event from the source machine, it updates the \texttt{ExtentCenter}, as illustrated in Figure~\ref{fig:modeled_en}.

In the modeled EN sync logic, the machine is driven by an external timer modeled in \psharp (see \S\ref{sec:method:timer}). It prepares a sync report with \texttt{extCtr.GetSyncReport(...)}, and then asynchronously sends the report to \texttt{ExtentManager} using \texttt{PSharp.Send(...)}. The \texttt{ExtentNode} machine also includes failure-related logic (see \S\ref{sec:method:driver}).

@@

  [OnEvent(RepeatedEvent, GenerateTimerTick)]
  void GenerateTimerTick() {
    // nondeterministic choice controlled by P#
    if (PSharp.Nondet())
      PSharp.Send(Target, new TimerTickEvent());
    PSharp.Send(this, new RepeatedEvent()); // loop
  }
}
\end{lstlisting}
\vspace{-4mm}
\caption{Modeling timer expiration using \psharp.}
\label{fig:modeled_timer}
\vspace{-2mm}
\end{figure}

@@

Figure~\ref{fig:modeled_timer} shows the \texttt{Timer} machine in the test harness. \texttt{Timer} invokes the \psharp method \texttt{Nondet()}, which generates a nondeterministic choice controlled by the \psharp runtime. Using \texttt{Nondet()} allows the machine to nondeterministically send a timeout event to its target (the \texttt{ExtentManager} or \texttt{ExtentNode} machines). The \psharp testing engine has the freedom to schedule arbitrary interleavings between these timeout events and all other regular system events.

@@

  void InjectNodeFailure() {
    // nondeterministically choose an EN using P#
    var node = (Machine)PSharp.Nondet(ExtentNodes);    
    PSharp.Send(node, new FailureEvent()); // fail chosen EN
  }
}
\end{lstlisting}
\vspace{-4mm}
\caption{Code snippet of the TestingDriver machine.}
\label{fig:testing_driver}
\vspace{-2mm}
\end{figure}

@@

Figure~\ref{fig:testing_driver} illustrates how \texttt{TestingDriver} leverages \psharp to inject nondeterministic failures. It uses \texttt{Nondet()} to nondeterministically choose an \texttt{ExtentNode} machine, and then sends a \texttt{FailureEvent} to the chosen machine to emulate an EN failure. As shown in the earlier Figure~\ref{fig:modeled_en}, the chosen \texttt{ExtentNode} machine processes the \texttt{FailureEvent}, notifies the liveness monitor of its failure (see \S\ref{sec:method:monitor}) and terminates itself by invoking the \psharp method \texttt{Halt()}. \psharp not only enumerates interleavings of asynchronous event handlers, but also the values returned by calls to \texttt{Nondet()}, thus enumerating different failure scenarios.

@@

\texttt{RepairMonitor} is a \psharp liveness monitor (see \S\ref{sec:overview:liveness}) that transitions between a cold and a hot state. Whenever an EN fails, the monitor is notified with an \texttt{ENFailedEvent} event. As soon as the number of extent replicas falls below a specified target (three replicas in the current \psharp test harness), the monitor transitions into the hot \emph{repairing} state, waiting for all missing replicas to be repaired. Whenever an extent replica is repaired, \texttt{RepairMonitor} is notified with an \texttt{ExtentRepairedEvent} event. When the replica number reaches again the target, the monitor transitions into the cold \emph{repaired} state, as illustrated in the code snippet of Figure~\ref{fig:monitor}.

@@

\subsection{Liveness Bug in Azure Storage vNext}
\label{sec:method:azurestore}
